---
title: "Games Summary"
date: 'Last updated: `r format(Sys.Date())`'
output:
  html_document:
    dpi: 500
    fig_width: 10
    fig_height: 7
---

```{r, echo=FALSE}

# Insert NFL logo in top right of report
htmltools::img(src = knitr::image_uri("NFL.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width: 110px; height: 128px')

```


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```



```{r Setup , echo =FALSE}

library(nflfastR)
library(nflseedR)
library(tidyverse)
library(teamcolors) # NFL team colors and logos
library(extrafont) # for extra fonts
library(ggrepel) # better labels
library(ggimage)
library(glue)
library(ggtext)
library(reactable)
library(reactablefmtr)
library(gt)
library(gtExtras)
library(ggforce)
library(ggsci)
library(ggimage)
library(animation)
library(DBI)
library(RSQLite)

# Optional but makes R prefer not to display numbers in scientific notation
options(scipen = 9999)

#Load games data from Lee Sharpe/nflseedR
games <- load_sharpe_games() %>% 
  filter(season == 2026)

```


```{r, echo=FALSE}

# Define 538 table theme for Reactable table(s) below
theme_538 <- function() {
    reactable::reactableTheme(
        searchInputStyle = list(width = "31%", backgroundColor = "#F9F9F9"),
        style = list(
            fontFamily = "Titillium Web"
        ),
        headerStyle = list(
            "&:hover[aria-sort]" = list(
                background = "hsl(0, 0%, 80%)"),
            "&[aria-sort='ascending'], &[aria-sort='descending']" = list(
                background = "#555",
                color = "#FFF"
            ),
            borderColor = "#333"
        ),
        borderColor = "#CDCDCD"
    )
}
```


## Games {.tabset}

### Scores & Schedule

```{r, echo=FALSE}

# Make games table
games2 <- games %>% 
  # using PFR's methodology to turn spreads into win probabilities
   mutate(home_pred = 1- (pnorm(0.5, mean = spread_line, sd = 12.7)
                    ),
          away_pred = 1-home_pred
         ) %>%
  select(game_id, game_type, gameday, week, away_team:home_score, spread_line, 
         away_pred, home_pred,away_qb_name, home_qb_name, stadium)

# Get team logos
team_logos <- nflfastR::teams_colors_logos

# Get player headshots
rosters <- nflfastR::fast_scraper_roster(seasons = 2023) %>% 
  select(full_name, headshot_url)


# join in team logos and headshots for inclusion in table
games2 <- left_join(games2, team_logos, by = c("away_team" = "team_abbr")) #Away teams
games2 <- left_join(games2, team_logos, by = c("home_team" = "team_abbr")) #Home teams

#games2 <- left_join(games2, rosters, by = c("away_qb_name" = "full_name")) #Away QBs
#games2 <- left_join(games2, rosters, by = c("home_qb_name" = "full_name")) #Home QBs


# set it up for the table
games2 <- games2 %>% 
  select(week, gameday, game_type, team_logo_espn.x, away_team, away_score, team_logo_espn.y, home_team, home_score, spread_line,  away_pred, home_pred) %>% 
  rename(team_logo_espn_away = team_logo_espn.x,
         team_logo_espn_home = team_logo_espn.y)
# %>% drop_na() #uncomment once first game has been played


# Reactable table

games_table <- reactable(
  games2,
  theme = theme_538,
  columnGroups = list(
                      colGroup(name = "Pre-Game Vegas Odds", 
                               columns = c("spread_line", "away_pred", "home_pred"))
                  ),
                  showSortIcon = TRUE,
                  searchable = TRUE,
                  language = reactableLang(
                      searchPlaceholder = "SEARCH FOR A TEAM (ABBR)..."),
                  defaultPageSize = 100,
                  columns = list(
                      week = colDef(name = "Week",
                                       align = "center"),
                      gameday = colDef(name = "Date",
                                       align = "right",
                                       minWidth = 120),
                      game_type = colDef(name = "Round",
                                         style = list(borderRight = "2px solid #555"),
                                       align = "right"),
                      away_team = colDef(name = "Away",
                                       align = "right"),
                      away_score = colDef(
                        name = "Away Score",
                        style = if (all(is.na(games2$away_score))) {
                          NULL  # Don't apply color scales if all values are NA
                        } else {
                          color_scales(
                            games2,
                            colors = paletteer::paletteer_d(palette = "ggsci::amber_material")
                          )
                        },
                        align = "right"
                      ),
                    home_team = colDef(name = "Home",
                                       align = "right"),
                    home_score = colDef(
                      name = "Home Score",
                      style = if (all(is.na(games2$home_score))) {
                        NULL  # Don't apply color scales if all values are NA
                      } else {
                        color_scales(
                          games2,
                          colors = paletteer::paletteer_d(palette = "ggsci::amber_material")
                        )
                      },
                      align = "right"
                    ),
                      spread_line = colDef(name = "Home Spread",
                                           style = list(borderLeft = "2px solid #555"),
                                       align = "right"),
                      away_pred = colDef(name = "Away Prob",
                                         format =  colFormat(percent = TRUE, digits = 0),
                                       align = "right"),
                      home_pred = colDef(name = "Home Prob",
                                         format =  colFormat(percent = TRUE, digits = 0),
                                       align = "right"),
                      
                      ### add logos using embed_img()
                      team_logo_espn_home = colDef(
                          name = "",
                          maxWidth = 40,
                          align = "right",
                          cell = embed_img(height = "25", width = "30")
                      ),
                      team_logo_espn_away = colDef(
                          name = "",
                          maxWidth = 40,
                          align = "right",
                          cell = embed_img(height = "25", width = "30")
                      )),
  
                  pagination = FALSE,
                  compact = TRUE, 
                  borderless = FALSE, 
                  striped = FALSE,
                  fullWidth = FALSE, 
                  defaultColDef = colDef(align = "center", minWidth = 95)
        ) %>% 
  add_title("2026 NFL Season",
            font_size = 28) %>% 
  add_subtitle("Every NFL game by week, including scores, schedules, and pre-game Vegas odds. Click a column header to sort by that statistic.",
               font_size = 14,
               font_weight = "normal") %>% 
  add_source("Data: nflverse")

games_table


```

### Game Recaps

A team-by-team box score for every game played this season, on the efficiency
metrics that actually explain results. Most recent game first, down to the
season opener in the last row, with each game a pair of rows and the away team
on top. Cells are shaded within the column: cool = good, warm = bad, so a row
that reads blue across is a team that controlled the game.

```{r Recap Data, echo = FALSE}

pbp <- load_pbp(2026)

# Run/pass plays with EPA, same filter as the League page
pbp_rp <- pbp %>%
  filter(rush == 1 | pass == 1, !is.na(epa))

current_week <- max(pbp_rp$week, na.rm = TRUE)

# Offensive and defensive EPA/play by team and game. Same definitions as the
# League page, where net EPA/play is off_epa - def_epa.
recap_off <- pbp_rp %>%
  group_by(game_id, week, team = posteam) %>%
  summarize(off_epa = mean(epa), .groups = "drop")

recap_def <- pbp_rp %>%
  group_by(game_id, team = defteam) %>%
  summarize(def_epa = mean(epa), .groups = "drop")

# Explosive plays: 20+ yards on passes, 10+ on runs (same as the Teams page).
recap_explosive <- pbp %>%
  filter(play_type %in% c("pass", "run"), penalty == 0, !is.na(epa)) %>%
  mutate(explosive_play = case_when(
    play_type == "pass" & yards_gained >= 20 ~ 1,
    play_type == "run" & yards_gained >= 10 ~ 1,
    TRUE ~ 0
  )) %>%
  group_by(game_id, team = posteam) %>%
  summarize(explosive_share = sum(explosive_play) / n(), .groups = "drop")

recap_turnovers <- pbp_rp %>%
  group_by(game_id, team = posteam) %>%
  summarize(
    n_turnovers = sum(interception, na.rm = TRUE) + sum(fumble_lost, na.rm = TRUE),
    turnover_rate = n_turnovers / n(),
    .groups = "drop"
  )

# Penalties are charged to penalty_team, which is not always posteam (defensive
# penalties happen on the other team's snaps), so this cannot come off pbp_rp.
recap_penalties <- pbp %>%
  filter(penalty == 1, !is.na(penalty_team), penalty_team != "") %>%
  group_by(game_id, team = penalty_team) %>%
  summarize(
    n_penalties = n(),
    penalty_yards = sum(penalty_yards, na.rm = TRUE),
    .groups = "drop"
  )

recaps <- recap_off %>%
  left_join(recap_def, by = c("game_id", "team")) %>%
  left_join(recap_explosive, by = c("game_id", "team")) %>%
  left_join(recap_turnovers, by = c("game_id", "team")) %>%
  left_join(recap_penalties, by = c("game_id", "team")) %>%
  mutate(
    across(c(n_turnovers, n_penalties, penalty_yards), ~ replace_na(.x, 0)),
    epa_play = off_epa - def_epa
  ) %>%
  left_join(
    games %>% select(game_id, gameday, gametime, away_team, home_team,
                     away_score, home_score),
    by = "game_id"
  ) %>%
  filter(!is.na(home_score)) %>%
  mutate(
    score = if_else(team == home_team, home_score, away_score),
    opp_score = if_else(team == home_team, away_score, home_score),
    home_away = if_else(team == home_team, "home", "away")
  ) %>%
  left_join(
    nflfastR::teams_colors_logos %>% select(team_abbr, team_logo_espn),
    by = c("team" = "team_abbr")
  ) %>%
  mutate(kickoff = as.POSIXct(paste(gameday, gametime), tz = "America/New_York")) %>%
  # Newest week on top, and within a week the latest kickoff first, so the last
  # row of the table is the season opener. game_id breaks the ties inside the
  # simultaneous Sunday windows, and home_away puts the away row above the home
  # row within each game.
  arrange(desc(week), desc(kickoff), game_id, home_away) %>%
  select(week, team_logo_espn, team, score, opp_score, epa_play, off_epa,
         def_epa, explosive_share, n_turnovers, turnover_rate, n_penalties,
         penalty_yards)

```

```{r Recap Table, echo = FALSE}

# Diverging fills: warm pole = bad, cool pole = good, neutral grey at the
# middle (the same two arms as the Series Success chart on the Teams page).
recap_pal <- c("#C25E45", "#E8A97E", "#F0EFEC", "#A9CBE3", "#6FB3DC")

epa_domain <- function(x) c(-1, 1) * max(abs(x), na.rm = TRUE)

recap_gt_data <- recaps %>%
  mutate(
    week_label = factor(paste("Week", week),
                        levels = paste("Week", sort(unique(week), decreasing = TRUE))),
    result = case_when(score > opp_score ~ "W", score < opp_score ~ "L", TRUE ~ "T"),
    turnovers_lab = sprintf("%d (%.1f%%)", n_turnovers, 100 * turnover_rate),
    penalties_lab = paste0(n_penalties, "/", penalty_yards),
    # each game is two rows; rule off the second one
    pair_end = row_number() %% 2 == 0
  ) %>%
  select(week_label, team_logo_espn, team, score, epa_play, off_epa, def_epa,
         explosive_share, turnovers_lab, penalties_lab,
         result, turnover_rate, penalty_yards, pair_end)

recap_table <- recap_gt_data %>%
  gt(groupname_col = "week_label") %>%
  gt_img_rows(columns = team_logo_espn, height = 22) %>%
  cols_label(
    team_logo_espn  = "",
    team            = "Team",
    score           = "",
    epa_play        = "EPA/Play",
    off_epa         = "Off EPA/Play",
    def_epa         = "Def EPA/Play",
    explosive_share = "Explosive Play Rate",
    turnovers_lab   = "Turnovers",
    penalties_lab   = "Penalties (n/yds)"
  ) %>%
  fmt_number(columns = c(epa_play, off_epa, def_epa), decimals = 2) %>%
  fmt_percent(columns = explosive_share, decimals = 1) %>%
  # One data_color() call per column so each is scaled on its own range. The
  # EPA columns get an explicit symmetric domain so that the palette's neutral
  # grey lands on exactly 0 EPA rather than on the midpoint of whatever range
  # the season happens to have produced. The turnover and penalty columns
  # print combined text, so they are shaded from a hidden numeric column via
  # target_columns.
  data_color(columns = epa_play, palette = recap_pal,
             domain = epa_domain(recap_gt_data$epa_play)) %>%
  data_color(columns = off_epa, palette = recap_pal,
             domain = epa_domain(recap_gt_data$off_epa)) %>%
  data_color(columns = def_epa, palette = rev(recap_pal),
             domain = epa_domain(recap_gt_data$def_epa)) %>%
  data_color(columns = explosive_share, palette = recap_pal) %>%
  data_color(columns = turnover_rate, target_columns = turnovers_lab,
             palette = rev(recap_pal)) %>%
  data_color(columns = penalty_yards, target_columns = penalties_lab,
             palette = rev(recap_pal)) %>%
  cols_hide(c(result, turnover_rate, penalty_yards, pair_end)) %>%
  cols_align(align = "left", columns = c(team, score)) %>%
  cols_align(align = "center", columns = c(epa_play, off_epa, def_epa,
                                           explosive_share, turnovers_lab,
                                           penalties_lab)) %>%
  # winner's score in green, loser's in red
  tab_style(style = cell_text(color = "#1A7A1A", weight = "bold"),
            locations = cells_body(columns = score, rows = result == "W")) %>%
  tab_style(style = cell_text(color = "#B3261E", weight = "bold"),
            locations = cells_body(columns = score, rows = result == "L")) %>%
  tab_style(style = cell_text(color = "#555555", weight = "bold"),
            locations = cells_body(columns = score, rows = result == "T")) %>%
  tab_style(style = cell_borders(sides = "right", color = "#555555", weight = px(2)),
            locations = cells_body(columns = score)) %>%
  tab_style(style = cell_borders(sides = "bottom", color = "#555555", weight = px(2)),
            locations = cells_body(rows = pair_end)) %>%
  tab_header(
    title = "Game Recaps",
    subtitle = glue("How each team performed in every game this season, through Week {current_week}. Cool cells are good, warm cells are bad, shaded within each column.")
  ) %>%
  tab_source_note("Data: nflfastR") %>%
  opt_all_caps(locations = "column_labels") %>%
  opt_table_font(font = list(google_font("Titillium Web"), default_fonts())) %>%
  tab_options(
    table.font.size = px(13),
    heading.title.font.size = px(28),
    heading.subtitle.font.size = px(14),
    heading.align = "left",
    heading.border.bottom.style = "none",
    column_labels.font.weight = "bold",
    column_labels.border.top.style = "none",
    row_group.font.weight = "bold",
    row_group.background.color = "#F9F9F9",
    table.border.top.style = "none",
    data_row.padding = px(3),
    source_notes.font.size = px(11)
  )

recap_table

```
