---
title: "Players Weekly Report"
date: 'Last updated: `r format(Sys.Date())`'
output:
  html_document:
    dpi: 500
    fig_width: 10
    fig_height: 7
---

```{r, echo=FALSE}

# Insert NFL logo in top right of report
htmltools::img(src = knitr::image_uri("/Users/Stephan/Desktop/R Projects/NFL/NFL.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width: 110px; height: 128px')

```


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

Below you'll find figures showing key advanced metrics I'm tracking for individual players throughout the season.

```{r, echo =FALSE}

library(nflfastR)
library(tidyverse)
library(teamcolors) # NFL team colors and logos
library(extrafont) # for extra fonts
library(ggrepel) # better labels
library(ggimage)
library(glue)
library(ggtext)
library(reactable)
library(reactablefmtr)
library(ggalt) #for dumbbell plot
library(ggforce)
library(gt)
library(gtExtras)

# Optional but makes R prefer not to display numbers in scientific notation
options(scipen = 9999)


data <- load_pbp(2024) #Load PBP data from nflfastR

# Filter for Run/Pass plays for the current season only
pbp_rp <- data %>%
  filter(rush == 1 | pass == 1, !is.na(epa)) # Exclude plays with missing EPA.

# Identify current week
current_week <- max(pbp_rp$week)

```

 
```{r, echo=FALSE}

library(gt)

# Create 538 GT table theme from Thomas Mock's blog.Comes from this post: https://themockup.blog/posts/2020-09-26-functions-and-themes-for-gt-tables/?panelset3=theme-code2&panelset4=theme-code3 
gt_nfl_theme_538 <- function(data,...) {
  data %>%
    # Add team logos w/ web_image
    text_transform(
      locations = cells_body(
        vars(team_logo_espn)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      }
    ) %>%
    # Relabel columns
    cols_label(
      team_logo_espn = ""
    ) %>%
    opt_all_caps()  %>%
    opt_table_font(
      font = list(
        google_font("Chivo"),
        default_fonts()
      )
    ) %>%
    tab_style(
      style = cell_borders(
        sides = "bottom", color = "transparent", weight = px(2)
      ),
      locations = cells_body(
        columns = TRUE,
        # This is a relatively sneaky way of changing the bottom border
        # Regardless of data size
        rows = nrow(data$`_data`)
      )
    )  %>% 
    tab_options(
      column_labels.background.color = "white",
      table.border.top.width = px(3),
      table.border.top.color = "transparent",
      table.border.bottom.color = "transparent",
      table.border.bottom.width = px(3),
      column_labels.border.top.width = px(3),
      column_labels.border.top.color = "transparent",
      column_labels.border.bottom.width = px(3),
      column_labels.border.bottom.color = "black",
      data_row.padding = px(3),
      source_notes.font.size = 12,
      table.font.size = 16,
      heading.align = "left",
      ...
    ) 
}

# Define 538 table theme for Reactable table(s) below
theme_538 <- function() {
    reactable::reactableTheme(
        searchInputStyle = list(width = "31%", backgroundColor = "#F9F9F9"),
        style = list(
            fontFamily = "Chivo"
        ),
        headerStyle = list(
            "&:hover[aria-sort]" = list(
                background = "hsl(0, 0%, 80%)"),
            "&[aria-sort='ascending'], &[aria-sort='descending']" = list(
                background = "#555",
                color = "#FFF"
            ),
            borderColor = "#333"
        ),
        borderColor = "#CDCDCD"
    )
}

# Custom ggplot theme (inspired by Owen Phillips at the F5 substack blog)
theme_custom <- function () { 
  theme_minimal(base_size=11, base_family="Chivo") %+replace% 
    theme(
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = 'floralwhite', color = "floralwhite")
    )
}

#Set aspect ratio for logo based plots
asp_ratio <- 1.618

```

***
## Quarterback Performance

### Quarterback Summary

The below shows a summary of QB performances so far this season.

```{r, echo=FALSE}

#load data
player_stats <- load_player_stats()

nfl_qb <- player_stats |>
  filter(position=="QB")

nfl_team <- teams_colors_logos

#reshape data
qb_summary_data <- nfl_qb |>
  filter(passing_yards>100)|> 
  select(headshot_url, player_name, player_display_name, recent_team, completions, attempts, passing_yards, passing_tds, interceptions, passing_epa, sacks, rushing_yards, rushing_tds)|>
  group_by(headshot_url, player_name, player_display_name, recent_team,) |> 
  summarise(total_completions = sum(completions),
            total_attempts = sum(attempts),
            total_passing_yards = sum(passing_yards),
            total_passing_tds = sum(passing_tds),
            total_interceptions = sum(interceptions),
            total_passing_epa = sum(passing_epa),
            total_sacks = sum(sacks),
            total_rushing_yards = sum(rushing_yards),
            total_rushing_tds = sum(rushing_tds)) |> 
  inner_join(teams_colors_logos|>
               select(team_abbr, team_nick, team_division, team_logo_espn), by=c("recent_team"="team_abbr"))|>
  separate(team_division, into=c("conference","division"), sep=" ")|>
  mutate(comp_perc = total_completions/total_attempts)|>
  arrange(conference, division, -total_passing_epa, -total_passing_tds, -total_passing_yards)|>
  group_by(conference, division)|>
  mutate(rank = row_number(), 
         pos = -rank)


# Create gt table
qb_summary_table <- qb_summary_data |> 
  select(headshot_url, player_display_name, team_logo_espn, recent_team:total_interceptions, conference:comp_perc, total_passing_epa) |> 
  gt() |> 
  tab_header(
    title = md("**Quarterback Passing Summary**"), 
    subtitle = glue("Minimum 100 pass yards, and ordered by passing TDs thrown. Thru Week {current_week}.")
    )  %>%
  cols_label(headshot_url = "",
             player_display_name = "Player",
             recent_team = "Team",
             team_logo_espn = "",
             total_completions = "Comps",
             total_attempts = "Att",
             total_passing_yards = "Pass Yds",
             total_passing_tds = "Pass TDs",
             total_interceptions = "INTs",
             comp_perc = "Comp %",
             total_passing_epa = "Pass EPA") %>% 
  gt_img_rows(headshot_url) %>%
  gt_img_rows(team_logo_espn) %>%
  data_color(
    columns = vars(total_passing_epa),
    colors = scales::col_numeric(
      palette = paletteer::paletteer_d(
        palette = "RColorBrewer::PRGn",
        direction  = 1
      ) %>% as.character(),
      domain = NULL, 
      na.color = "#00441BFF"
    )
  ) %>%
   fmt_percent(
    columns = comp_perc,
    decimals = 1
  ) |> 
  fmt_number(
    columns = total_passing_epa,
    decimals = 1
  ) |> 
  tab_options(
    table.background.color = "white",
    heading.title.font.size  = 28,
    heading.title.font.weight = 'bold',
    heading.subtitle.font.size = 14,
    table.font.names = "Chivo", 
    table.font.color = 'black',
    table.border.top.color = "transparent",
    footnotes.font.size = 12,
    source_notes.font.size = 12,
    data_row.padding = px(2), 
    footnotes.padding = px(1), 
  ) %>%
  tab_source_note(
    source_note = md("Table: @steodosescu | Data: nflfastR")
  )

qb_summary_table

```

***

### EPA per Dropback by Team

The below isolates Expected Points Added for pass plays only to get a sense of how each team's quarterback is performing on a per-dropback basis.

```{r, echo = FALSE}

epa_play <- pbp_rp %>% 
  filter(pass == 1, !is.na(posteam)) %>% 
  group_by(posteam) %>% 
  summarize(
    n = n(),
    epa_per_db = sum(epa, na.rm = TRUE) / n,
    cpoe_per_db = sum(cpoe, na.rm = TRUE) / n,
    success_rate = sum(epa, na.rm = TRUE) / n
  )
  

epa_logos <- epa_play %>%
  left_join(nflfastR::teams_colors_logos, by = c("posteam" = "team_abbr")) %>% 
  ggplot(aes(x = epa_per_db, y = reorder(posteam, epa_per_db))) +
  geom_image(aes(image = team_logo_espn), size = 0.045, by = "width", asp = asp_ratio) + 
  scale_fill_identity(aesthetics = c("fill", "colour")) +
  theme_custom() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.y = element_blank(),
  ) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(breaks = seq(-0.5, 0.5, 0.1)) +
  labs(
    x = "EPA per Dropback",
    y = "",
    title = "Team QB Efficiency, 2024 NFL Season",
    subtitle = glue("EPA per dropback measured thru **Week {current_week}**."),
    caption = "Data: @nflfastR | Plot: @steodosescu") +
  theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5)
    ) +
  theme(plot.subtitle = element_markdown(
    hjust = 0.5)
    )

epa_logos
```

<!-- ### EPA vs CPOE per dropback by team -->

<!-- The below is another way to look at quarterback efficiency using EPA and Completion Percentage Over Expectation (CPOE) as measures. CPOE is simply the difference between a quarterback’s expected completion percentage and actual completion percentage. Expected completion percentage is a stat measuring the likelihood of a given pass being completed which factors in features like depth of target (air yards). It's a better measure of accuracy than traditional completion percentage because it takes into account the location of where passes are being thrown. -->

<!-- ```{r EPA vs CPOE, echo = FALSE} -->

<!-- epa_cpoe_logos <- epa_play %>% -->
<!--   left_join(nflfastR::teams_colors_logos, by = c("posteam" = "team_abbr")) %>%  -->
<!--   ggplot(aes(x = cpoe_per_db, y = epa_per_db)) + -->
<!--   geom_image(aes(image = team_logo_espn), size = 0.065, by = "width", asp = asp_ratio) + -->
<!--   scale_fill_identity(aesthetics = c("fill", "colour")) + -->
<!--   theme_custom() + -->
<!--   geom_hline(yintercept = 0) + -->
<!--   labs( -->
<!--     x = "Completion % Over Expectation (CPOE)", -->
<!--     y = "Expected Points Added (EPA) per Dropback", -->
<!--     title = "Team QB Performance, 2023 NFL Season", -->
<!--     subtitle = glue("Perfomance as defined by a composite of EPA and CPOE. Thru **Week {current_week}**"), -->
<!--     caption = "Data: @nflfastR | Plot: @steodosescu") + -->
<!--   geom_hline(yintercept = mean(epa_play$epa_per_db), color = "red", linetype = "dashed") + -->
<!--   geom_vline(xintercept = mean(epa_play$cpoe_per_db), color = "red", linetype = "dashed") + -->
<!--   theme(plot.title = element_text(face = "bold",  -->
<!--                                   size = 20,  -->
<!--                                   hjust = 0.5 -->
<!--     )) + -->
<!--   theme(plot.subtitle = element_markdown( -->
<!--     hjust = 0.5) -->
<!--     ) -->

<!-- epa_cpoe_logos -->
<!-- ``` -->

### EPA vs CPOE per dropback by team

The below is another way to look at quarterback efficiency using EPA and Completion Percentage Over Expectation (CPOE) as measures. CPOE is simply the difference between a quarterback’s expected completion percentage and actual completion percentage. Expected completion percentage is a stat measuring the likelihood of a given pass being completed which factors in features like depth of target (air yards). It's a better measure of accuracy than traditional completion percentage because it takes into account the location of where passes are being thrown.

Here we focus on quarterback play only as the QB position is the most valuable position on the field. As the quarterback goes, the team goes.

```{r}

qbs <- pbp_rp %>%
  filter(
    play_type %in% c("pass", "run"),
    penalty == 0,
    !is.na(epa)
  ) %>% 
  group_by(name, posteam) %>%
  summarize(
    n_dropbacks = sum(pass),
    n_rush = sum(rush),
    n_plays = n(),
    epa_per_play = sum(epa) / n_plays,
    cpoe_per_play = sum(cpoe, na.rm = TRUE) / n_plays,
    success_per_play = sum(success) / n_plays
  ) %>%
  filter(n_dropbacks >= 40) %>% #Change this No. later in the season
  ungroup() # always ungroup if you no longer need the grouping effect


qb_success_rate <- qbs %>%
  left_join(nflfastR::teams_colors_logos, by = c("posteam" = "team_abbr")) %>%
  ggplot(aes(x = cpoe_per_play, y = epa_per_play)) +
  geom_image(aes(image = team_logo_espn), size = 0.030, by = "width", asp = asp_ratio) +
  # Notice that color/size inside aes()
  #geom_point(aes(color = if_else(posteam == "CHI", "#C83803", "#0B162A"), size = n_plays / 60), alpha = 0.50) +
  # we need this to assign red/black to the actual color
  scale_color_identity() +
  
  # add labels for all players
  geom_text_repel(aes(label = name, color = if_else(posteam == "CHI", "#C83803", "#0B162A")),
                  force = 1, point.padding = 0.1,
                  segment.size = 0.2
  ) +
  labs(
    x = "Completion % Over Expecation (CPOE)",
    y = "EPA per play",
    caption = "Data: @nflscrapR | Plot: @steodosescu",
    title = "Quarterback Performance",
    subtitle = "2024 season, min 40 pass attempts, includes all QB's rush and pass plays."
  ) +
  theme_custom() +
  theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    )) +
  theme(plot.subtitle = element_markdown(
    hjust = 0.5)
    ) +
  theme(legend.position = "none")

qb_success_rate +
  stat_smooth(method = "lm", geom = "line", alpha = 0.5, se = FALSE, color = "red", size = 1)

```

### Difference in Expected and Actual Yards

The below shows what quarterbacks have completed a either higher percentage or lower percentage of their passes than expected, according to the nflfastR model.

```{r, echo = FALSE}

qb_cpoe <- pbp_rp %>%
  filter(
    play_type %in% c("pass"),
    penalty == 0,
    !is.na(epa)
  ) %>% 
  group_by(name, posteam) %>%
  summarize(
    n_dropbacks = sum(pass),
    n_rush = sum(rush),
    n_plays = n(),
    epa_per_play = sum(epa) / n_plays,
    success_per_play = sum(success) / n_plays,
    cpoe_per_db = round(sum(cpoe, na.rm = TRUE) / n_plays,2),
    cp_per_db = round(sum(cp, na.rm = TRUE) / n_plays,2)
  ) %>%
  mutate(exp_cp_per_db = round(cp_per_db - cpoe_per_db/100,2)) %>% 
  mutate(pos_neg = case_when(
    cpoe_per_db >= 0 ~ "Overachieving", 
    TRUE ~ "Underachieving",
  )) %>% 
  filter(n_dropbacks >= 40) %>% #Change this No. later in the season
  ungroup() # always ungroup if you no longer need the grouping effect

# Make QB CPOE dumbbell plot (Commenting this out in favor of the comet plot below)

# qb_cpoe_plot <- qb_cpoe %>% 
  #  ggplot(aes(x = exp_cp_per_db, xend = cp_per_db, y = reorder(name, cpoe_per_db), group = name)) # + 
  #  geom_dumbbell(colour = "#dddddd",
   #               size = 2,
    #              colour_x = "#1380A1",
     #             colour_xend = "#FAAB18") + 
    #labs(x = "Rate (%)", y = "",
     #    title = "Who's Overachieving in the Pass Game?", 
      #   subtitle = glue("Difference between <span style = 'color:#1380A1'>Expected Completion # %</span> and <span style = 'color:#FAAB18'>Actual Completion %</span>. Minimum 20 passes."),
#         caption = "Data Source: nflfastR \nGraphic: @steodosescu") +
 #   theme_custom() +
  #  theme(plot.title = element_text(face="bold"), text = element_text(family = "Chivo")) +
   # theme(plot.subtitle = element_markdown()) +
    #geom_text(data=qb_cpoe, aes(x=cp_per_db, y=name, label=cp_per_db),
     #         color="#FAAB18", size=2.75, vjust=2.5, family="Chivo") +
    #geom_text(data=qb_cpoe, aes(x=exp_cp_per_db, y=name, label=exp_cp_per_db),
     #         color="#1380A1", size=2.75, vjust=2.5, family="Chivo") +
    # geom_rect(data=qb_cpoe, aes(xmin=0.82, xmax=0.88, ymin=-Inf, ymax=Inf), fill="lightgrey") +
    #geom_text(data=qb_cpoe, aes(label = cpoe_per_db, y=name, x=0.85), fontface="bold", size=3, # family="Chivo")

# qb_cpoe_plot


# Make comet plot

comet_plot <- qb_cpoe %>% 
  ggplot() + 
  geom_link(aes(x = exp_cp_per_db, y = reorder(name, cpoe_per_db), xend = cp_per_db, yend = reorder(name, cpoe_per_db), color = pos_neg, size = stat(index))) +
  scale_color_manual(values = c("#00A087FF", "#E64B35FF")) +
  scale_size(range = c(.01, 4)) + 
  # scale_x_continuous(labels = c("-10", "-5", "0", "+5", "+10"), breaks = seq(-10, 10, 5)) +
  theme_custom() +
  geom_point(
    data = filter(qb_cpoe, cpoe_per_db > 0),
    aes(cp_per_db, y = reorder(name, cpoe_per_db), color = pos_neg),
    shape = 21,
    fill = "white",
    size = 3.5
  )  +
  geom_point(
    data = filter(qb_cpoe, cpoe_per_db < 0),
    aes(cp_per_db, y = reorder(name, cpoe_per_db), color = pos_neg),
    shape = 21,
    fill = "white",
    size = 3.5
  ) + 
  theme(legend.position = 'none') +
  labs(x = "Rate (%)", y = "",
         title = "Who's Over/Underachieving in the Pass Game?", 
         subtitle = glue("Difference between **Expected Completion %** and **Actual Completion %**. Minimum 40 passes. Thru Week {current_week}"),
         caption = "Data Source: nflfastR \nGraphic: @steodosescu") +
    theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    )) +
    theme(plot.subtitle = element_markdown()) +
    geom_rect(data=qb_cpoe, aes(xmin=0.83, xmax=0.87, ymin=-Inf, ymax=Inf), fill="lightgrey") +
    geom_text(data=qb_cpoe, aes(label = cpoe_per_db, y=name, x=0.85), fontface="bold", size=3, family="Chivo") +
  annotate(geom = 'label', x = 0.50, y = 4.5, label = "Less Completions\nThan Expected", family = "Chivo", color = "#E64B35FF", fontface = 'bold', fill = "floralwhite", label.size = 0, size = 3) +
  annotate(geom = 'label', x = 0.75, y = 24.5, label = "More Completions\nThan Expected", family = "Chivo", color = "#00A087FF", fontface = 'bold', fill = "floralwhite", label.size = 0, size = 3)

comet_plot


```



***
## Receivers Performance

### Receivers average Depth of Target (aDOT) vs. Catch Rate

The below shows receiver catch rate against their average depth of target (air yards).

```{r, echo = FALSE}

receiving <- pbp_rp %>%
  filter(!is.na(receiver_player_name) & play_type == "pass" & down <= 4) %>%
  group_by(receiver_player_name) %>%
  summarize(adot = mean(air_yards), targets = n(), catch_rate = sum(complete_pass)/targets) %>%
  filter(targets >= 10) %>% #Change this No. later in the season
  ungroup() # always ungroup if you no longer need the grouping effect


# Make plot

receiver_plot <- receiving %>% 
  ggplot(aes(x = adot, y = catch_rate)) + 
  geom_point(size = receiving$targets/10, alpha = 0.5) + 
  geom_text_repel(aes(label = receiver_player_name), color = "grey40", segment.color = "black") + 
  geom_smooth(method = lm, alpha = 0.5, linetype = "dashed", se = FALSE, color = "red", size = 1) + 
  scale_y_continuous(breaks = c(.4,.45,.5,.55,.6,.65,.7,.75,.8,.85)) +
  theme_custom() + 
  labs(x = "aDOT (Yards)", 
       y = "Catch Rate (%)",
       title = "Catch Rate vs aDOT",
       subtitle = glue("Minimum 10 Targets. Average Depth of Target is based on Air Yards. Thru Week **{current_week}**."),
       caption = "Data Source: nflfastR \nGraphic: @steodosescu"
       ) +
      theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    )) +
    theme(plot.subtitle = element_markdown(
    hjust = 0.5)
    )

receiver_plot
  
```

***
## Rushing Performance

The below shows a rusher's success rate vs his total EPA gained on runs this season.

```{r, echo = FALSE}

rushing <- pbp_rp %>%
  filter(!is.na(rusher_player_name) & play_type == "run" & down <= 4) %>%
  group_by(rusher_player_name, posteam) %>%
  summarize(n_rush=sum(rush),
            epa_per_rush=sum(epa*rush)/n_rush,
            #epa = sum(epa, na.rm = TRUE) / n,
            success_per_rush=sum(rush*epa>0)/n_rush) %>%
  filter(n_rush >= 20
         ) %>% #Change this No. later in the season
  ungroup() # always ungroup if you no longer need the grouping effect


# Make plot

rushing_plot <- rushing %>% 
  left_join(nflfastR::teams_colors_logos, by = c("posteam" = "team_abbr")) %>%
  ggplot(aes(x = success_per_rush, y = epa_per_rush)) + 
  geom_image(aes(image = team_logo_espn), size = 0.030, by = "width", asp = asp_ratio) +
  scale_color_identity() +
  
  #ggplot(aes(x = success_per_rush, y = epa_per_rush)) + 
  #geom_point(size = receiving$targets/20, alpha = 0.5) + 
  # add labels for all players
  geom_text_repel(aes(label = rusher_player_name), segment.color = "black") + 
  geom_smooth(method = lm, alpha = 0.5, linetype = "dashed", se = FALSE, color = "red", size = 1) + 
  #scale_y_continuous(breaks = c(.4,.45,.5,.55,.6,.65,.7,.75,.8,.85)) +
  theme_custom() + 
  labs(x = "Success Rate (%)", 
       y = "EPA/Rush",
       title = "Ground Attack: Rushing Efficiency",
       subtitle = glue("Minimum 20 Rushes. Thru Week **{current_week}**."),
       caption = "Data Source: nflfastR \nGraphic: @steodosescu"
       ) +
      theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    )) +
    theme(plot.subtitle = element_markdown(
    hjust = 0.5)
    )

rushing_plot
  
```