---
title: "Teams Weekly Report"
date: 'Last updated: `r format(Sys.Date())`'
output:
  html_document:
    dpi: 500
    fig_width: 10
    fig_height: 7
---

```{r, echo=FALSE}

# Insert NFL logo in top right of report
htmltools::img(src = knitr::image_uri("NFL.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width: 110px; height: 128px')

```


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

Below you'll find figures showing key advanced metrics I'm tracking for all 32 NFL teams.

```{r Setup , echo =FALSE}

library(nflfastR)
library(nflseedR)
library(tidyverse)
library(teamcolors) # NFL team colors and logos
library(extrafont) # for extra fonts
library(ggrepel) # better labels
library(ggimage)
library(glue)
library(ggtext)
library(nflplotR) # team logos on plot axes
library(reactable)
library(reactablefmtr)
library(ggforce)
library(ggsci)
library(paletteer)
library(prismatic)

# load pbp data
#data <- load_pbp(2013:2024)
data <- load_pbp(2026)

# Filter for Run/Pass plays for the current season only
pbp_rp <- data %>%
  filter(season == 2026) %>%
  filter(rush == 1 | pass == 1, !is.na(epa)) # Exclude plays with missing EPA.

# Identify current week
current_week <- max(pbp_rp$week)

```

 
```{r Set Themes, echo=FALSE}

library(gt)
library(gtExtras)

# Create 538 GT table theme from Thomas Mock's blog.Comes from this post: https://themockup.blog/posts/2020-09-26-functions-and-themes-for-gt-tables/?panelset3=theme-code2&panelset4=theme-code3 
gt_nfl_theme_538 <- function(data,...) {
  data %>%
    # Add team logos w/ web_image
    text_transform(
      locations = cells_body(
        vars(team_logo_espn)
      ),
      fn = function(x) {
        web_image(
          url = x,
          height = 25
        )
      }
    ) %>%
    # Relabel columns
    cols_label(
      team_logo_espn = ""
    ) %>%
    opt_all_caps()  %>%
    opt_table_font(
      font = list(
        google_font("Titillium Web"),
        default_fonts()
      )
    ) %>%
    tab_style(
      style = cell_borders(
        sides = "bottom", color = "transparent", weight = px(2)
      ),
      locations = cells_body(
        columns = TRUE,
        # This is a relatively sneaky way of changing the bottom border
        # Regardless of data size
        rows = nrow(data$`_data`)
      )
    )  %>% 
    tab_options(
      column_labels.background.color = "white",
      table.border.top.width = px(3),
      table.border.top.color = "transparent",
      table.border.bottom.color = "transparent",
      table.border.bottom.width = px(3),
      column_labels.border.top.width = px(3),
      column_labels.border.top.color = "transparent",
      column_labels.border.bottom.width = px(3),
      column_labels.border.bottom.color = "black",
      data_row.padding = px(3),
      source_notes.font.size = 12,
      table.font.size = 16,
      heading.align = "left",
      ...
    ) 
}

# Define 538 table theme for Reactable table(s) below
theme_538 <- function() {
    reactable::reactableTheme(
        searchInputStyle = list(width = "31%", backgroundColor = "#F9F9F9"),
        style = list(
            fontFamily = "Titillium Web"
        ),
        headerStyle = list(
            "&:hover[aria-sort]" = list(
                background = "hsl(0, 0%, 80%)"),
            "&[aria-sort='ascending'], &[aria-sort='descending']" = list(
                background = "#555",
                color = "#FFF"
            ),
            borderColor = "#333"
        ),
        borderColor = "#CDCDCD"
    )
}

# Custom ggplot theme (inspired by Owen Phillips at the F5 substack blog)
# Built on theme_bw and stripped back to match theme_minimal's look. Under
# ggplot2 4.0 a theme_minimal base drops custom axis-text elements, so logo
# axes (nflplotR::element_nfl_logo) render blank; other base themes are fine.
theme_custom <- function () {
  theme_bw(base_size=11, base_family="Titillium Web") %+replace%
    theme(
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      axis.ticks = element_blank(),
      legend.background = element_blank(),
      legend.key = element_blank(),
      strip.background = element_blank(),
      plot.background = element_rect(fill = 'floralwhite', color = "floralwhite"),
      plot.subtitle = element_text(color = "grey60")
    )
}


#Set aspect ratio for logo based plots
asp_ratio <- 1.618

```

***
## Team Level Performance


### Offensive vs. Defensive EPA

Expected Points Added: The change in Expected Points (EP) from one play to the next. EPA yields a single measure of the value of every play. Expected Points was created by the nflfastR team using a statistical model trained on historical data, and takes into account features like down, distance, time remaining, whether the game is being played indoors, etc. It helps answer the question “How good of a position is my team in to score as of now?” Higher EPA on offense is better, while a lower EPA on defense is more desirable. 

EPA helps provide better context around what plays are more valuable. For example, a five-yard completion on third-and-4 is better than an eight-yard completion on third-and-9, despite the fact that the latter resulted in more yards. 

```{r}

offense <- pbp_rp %>%
    group_by(posteam)%>%
    summarize(
        n_pass=sum(pass),
        n_rush=sum(rush),
        epa_per_pass=sum(epa*pass)/n_pass,
        epa_per_rush=sum(epa*rush)/n_rush,
        success_per_pass=sum(pass*epa>0)/n_pass,
        success_per_rush=sum(rush*epa>0)/n_rush,
        off_epa=mean(epa))

defense <- pbp_rp %>%
    group_by(defteam)%>%
    summarize(
        def_n_pass=sum(pass),
        def_n_rush=sum(rush),
        def_epa_per_pass=sum(epa*pass)/def_n_pass,
        def_epa_per_rush=sum(epa*rush)/def_n_rush,
        def_success_per_pass=sum(pass*epa>0)/def_n_pass,
        def_success_per_rush=sum(rush*epa>0)/def_n_rush,
        def_epa=mean(epa)
    )

off_def_epa <- offense %>% 
    inner_join(defense, by=c("posteam" = "defteam")) %>%
    left_join(nflfastR::teams_colors_logos, by = c("posteam" = "team_abbr"))


# Make EPA Tiers plot

epa_plot <- off_def_epa %>% 
    ggplot(aes(x = off_epa, y = def_epa)) + 
    geom_image(aes(image = team_logo_espn), size = 0.065, by = "width", asp = asp_ratio) +
    geom_hline(yintercept = mean(off_def_epa$off_epa), color = "red", linetype = "dashed") +
    geom_vline(xintercept =  mean(off_def_epa$def_epa), color = "red", linetype = "dashed") +
    theme_custom() +
  geom_abline(slope = -1.5, intercept = c(.4, .3, .2, .1, 0, -.1, -.2, -.3, -.4), alpha = .2) +
    labs(x = "Offense EPA/play",
         y = "Defense EPA/play",
         caption = "Data: @nflscrapR | Plot: @steodosescu",
         title = glue("Expected Points Added (EPA) Tiers"),
        subtitle = glue("Rush and pass plays only | All downs | All Quarters | As of Week {current_week}.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    ),
    plot.subtitle = element_text(
      size = 10,
      hjust = 0.5)
  ) +
  theme(plot.subtitle = element_markdown()) +
  scale_y_reverse()

epa_plot

```


### Offensive EPA

Expected Points Added for offensive units. Top right quadrant houses the units with the best offense.

```{r, echo = FALSE}

# Offensive EPA Plot
off_epa_plot <- off_def_epa %>% 
    ggplot(aes(x = epa_per_pass, y = epa_per_rush)) + 
    geom_image(aes(image = team_logo_espn), size = 0.065, by = "width", asp = asp_ratio) +
    geom_hline(yintercept = mean(off_def_epa$epa_per_pass), color = "red", linetype = "dashed") +
    geom_vline(xintercept =  mean(off_def_epa$epa_per_rush), color = "red", linetype = "dashed") +
    theme_custom() +
    labs(x = "EPA/pass",
         y = "EPA/rush",
         caption = "Data: @nflscrapR | Plot: @steodosescu",
         title = glue("Offensive EPA"),
        subtitle = glue("Rush and pass plays only | All downs | All Quarters | As of Week {current_week}.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    ),
    plot.subtitle = element_text(
      size = 10,
      hjust = 0.5)
  ) +
  theme(plot.subtitle = element_markdown())

off_epa_plot

```


### Defensive EPA

Expected Points Added for defensive units. Top right quadrant houses the units with the best defense.

```{r, echo = FALSE}

# Defensive EPA Plot
def_epa_plot <- off_def_epa %>% 
    ggplot(aes(x = def_epa_per_pass, y = def_epa_per_rush)) + 
    geom_image(aes(image = team_logo_espn), size = 0.065, by = "width", asp = asp_ratio) +
    geom_hline(yintercept = mean(off_def_epa$def_epa_per_pass), color = "red", linetype = "dashed") +
    geom_vline(xintercept =  mean(off_def_epa$def_epa_per_rush), color = "red", linetype = "dashed") +
    theme_custom() +
    labs(x = "EPA/pass Allowed",
         y = "EPA/rush Allowed",
         caption = "Data: @nflscrapR | Plot: @steodosescu",
         title = glue("Defensive EPA"),
        subtitle = glue("Rush and pass plays only | All downs | All Quarters | As of Week {current_week}.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    ),
    plot.subtitle = element_text(
      size = 10,
      hjust = 0.5)
  ) +
  theme(plot.subtitle = element_markdown()) +
    scale_x_reverse() +
    scale_y_reverse() 

def_epa_plot


```


***

### Performance vs Expectations

The below shows current win totals -- expressed in win percentage -- compared to pre-season expectations, according to Vegas bookmakers. Data was pulled from pro-football-reference.com the day before the regular season began for this analysis.

```{r, echo = FALSE}

# get current wins
games_df <- nflreadr::load_schedules() %>% 
  filter(season == 2026, game_type == "REG") %>% 
  select(game_id, team_home = home_team, team_away = away_team, result, week) %>% 
  pivot_longer(contains('team'), names_to = 'home_away', values_to = 'team', names_prefix = 'team_') %>% 
  mutate(
    result = ifelse(home_away == 'home', result, -result),
    win = ifelse(result == 0 , 0.5, ifelse(result > 0, 1, 0))
  ) %>% 
  select(week, team, win, result) %>% 
  drop_na()


team_df <- nflreadr::load_teams() %>% 
  select(team_logo_espn, team_abbr, team_name, team_conf, team_division, team_color)

games_df <- games_df %>% 
  group_by(team) %>% 
  summarise(
    Wins = length(win[win==1]),
    Losses = length(win[win==0]),
    result = sum(result)) %>%
  left_join(team_df, by = c("team" = "team_abbr")) %>% 
  select(team_logo_espn, team, team_name, team_conf, team_division, team_color, Wins, Losses, result)


# load preseason win totals from Vegas sportsbooks (accessed in late August)
vegas_totals <- read_csv("https://raw.githubusercontent.com/steodose/NFL/refs/heads/master/preseason_win_totals_2026.csv")

vegas_totals <- vegas_totals %>% 
    mutate(implied_win_perc = win_total/17)

# join vegas win totals to current win totals
joined_df <- games_df %>%
    left_join(vegas_totals, by = c("team" = "team")) %>%
  mutate(total_games = Wins + Losses,
         win_total_perc = Wins/total_games,
         win_total_diff = win_total_perc - implied_win_perc)

# create barplot
expected_totals_barplot <- joined_df %>% 
  ggplot(aes(x = fct_reorder(team_name, -win_total_diff), y = win_total_diff)) +
  geom_col(aes(fill = team_color, 
               color = after_scale(clr_darken(fill, 0.3))
  ),
  width = 0.4, 
  alpha = .75,
  ) + 
  scale_color_identity(aesthetics =  c("fill"))  +
  geom_image(
    aes(
      image = team_logo_espn                                  
    ), 
    size = 0.035, 
    by = "width", 
    asp = asp_ratio
  ) +
  geom_hline(yintercept = 0, color = "black", size = 1) +
  scale_y_continuous(labels = scales::percent_format(),
                       breaks = scales::pretty_breaks(n = 10)) +
  theme_custom() + 
  theme(axis.text.x = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(x = "", 
       y = "Win % over Expected", 
       caption = "Data: nflverse/OddsAPI | Plot: @steodosescu",
       title = glue("Performance Relative to Expecations"),
       subtitle = glue("Actual win totals vs. Predicted. Win total over/unders pulled from various sportsbooks as of September, 2026.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    ),
    plot.subtitle = element_text(
      hjust = 0.5)
  )
  
expected_totals_barplot


```


***
### Pythagorean Wins

The below shows each team's actual win percentage vs their pythagoroean win percentage. Pythagorean winning percentage is important to look at because it's a better predictor of future success than actual wins. According to Sumer Sports, pythagorean winning percentage has an R^2 of 0.14 to next year's winning percentage while actual winning percentage is only 0.11.

``` {r, echo = FALSE}

games_df2 <-  nflreadr::load_schedules() %>% 
  filter(season == 2026, game_type == "REG") %>% 
  select(game_id, team_home = home_team, team_away = away_team, away_score, home_score, result, week) %>% 
  drop_na() %>%
  pivot_longer(contains('team'), names_to = 'home_away', values_to = 'team', names_prefix = 'team_') %>% 
  mutate(
    result = ifelse(home_away == 'home', result, -result),
    win = ifelse(result == 0 , 0.5, ifelse(result > 0, 1, 0))
  ) %>% 
  mutate(
    PF = ifelse(home_away == 'home', home_score, away_score),
    PA = ifelse(home_away == 'home', away_score, home_score),
  ) %>% 
  group_by(team) %>%
  summarise(
    wins = length(win[win==1]),
    losses = length(win[win==0]),
    ties = length(win[win==0.5]),
    pf = sum(PF),
    pa = sum(PA))


# pyth_wins_df <- games_df2 %>%
#   mutate(total_games = wins + losses,
#          win_total_perc = wins/total_games,
#          pyth_win_perc = ((pf/pa)^2)/(1+(pf/pa)^2)) %>%
#   left_join(team_df, by = c("team" = "team_abbr")) 

pyth_wins_df <- games_df2 %>%
  mutate(total_games = wins + losses,
         win_total_perc = (wins/total_games)+(ties*0.005),
         pyth_win_perc = ((pf/pa)^2)/(1+(pf/pa)^2),
         pyth_wins = (pf^2.37/(pf^2.37+pa^2.37))*17) %>%
  left_join(team_df, by = c("team" = "team_abbr")) 


py_wins_plot <- pyth_wins_df %>% 
  ggplot(aes(x = win_total_perc, y = pyth_win_perc)) + 
  geom_image(aes(image = team_logo_espn), size = 0.065, by = "width", asp = asp_ratio) +
  annotate("text", x=0.3, y=0.65, label = "Better than Record", color = "#00BA38", family = 'Outfit') +
  annotate("text", x=0.7, y=0.35, label = "Worse than Record", color = "#F8766D", family = 'Outfit') +
  theme_custom() +
  #geom_abline(slope = 1, alpha = .2) +
  geom_smooth(method = "lm", se = FALSE, color="gray", alpha = 0.2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(x = "Actual Win %",
       y = "Pythagorean Win %",
       caption = "Data: @nflfastR | Plot: @steodosescu",
       title = glue("Actual and Pythagorean Win Percentage"),
       subtitle = glue("Pythagorean Win % tells how many games a team should have won based on points scored/allowed. It's an indicator of potential future success. As of Week {current_week}.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
        ),
        plot.subtitle = element_text(
          size = 10,
          hjust = 0.5)
  ) +
  theme(plot.subtitle = element_markdown())

py_wins_plot


```


***
### Point Differential

The below shows cumulative point differential for each team this season.

```{r, echo = FALSE}

# create point differential barplot
point_diff_barplot <- games_df %>% 
  ggplot(aes(x = fct_reorder(team_name, -result), y = result)) +
  geom_col(aes(fill = team_color, 
               color = after_scale(clr_darken(fill, 0.3))
  ),
  width = 0.4, 
  alpha = .75,
  ) + 
  scale_color_identity(aesthetics =  c("fill"))  +
  geom_image(
    aes(
      image = team_logo_espn                                  
    ), 
    size = 0.035, 
    by = "width", 
    asp = asp_ratio
  ) +
  geom_hline(yintercept = 0, color = "black", size = 1) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_custom() + 
  theme(axis.text.x = element_blank(), 
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank()) + 
  labs(x = "", 
       y = "Point Differential", 
       caption = "Data: nflverse | Plot: @steodosescu",
       title = glue("NFL Point Differential"),
       subtitle = glue("As of Week {current_week}.")) +
  theme(plot.title.position = "plot",
        plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    ),
    plot.subtitle = element_text(
      hjust = 0.5)
  )
  

point_diff_barplot

```


***
### Explosive Plays

Some teams are better than others at getting chunk plays. We're defining explosive plays as those gaining 15 yards or more on rushes and 20 yards+ on passes. We're highlighting the 49ers this season since they have been among the most exciting teams to watch. 

```{r Explosive Plays, echo = FALSE, fig.height = 10}

explosive_plays <- pbp_rp %>% 
  filter(
    play_type %in% c("pass", "run"),
    penalty == 0,
    !is.na(epa)
  ) %>% 
  mutate(explosive_play = case_when((play_type == "pass" & yards_gained >= 20) ~ 1,
                            (play_type == "run" & yards_gained >= 10) ~ 1,
                            TRUE ~ 0)) %>% 
  group_by(posteam) %>%
  summarize(
    n_dropbacks = sum(pass),
    n_rush = sum(rush),
    n_plays = n(),
    n_explosive_plays = sum(explosive_play)
  ) %>%
  mutate(explosive_share = n_explosive_plays/n_plays) %>% 
  ungroup() # always ungroup if you no longer need the grouping effect


# Helmets instead of team names on the y axis. nflplotR::element_path()
# renders axis text that is a file path as an image, so the axis values are
# the helmet paths themselves and the ordering rides along on the factor.
# (This replaced element_nfl_logo(), which in turn replaced <img> tags drawn
# with ggtext::element_markdown() -- that stopped working under ggplot2 4.0.)

explosive_plays <- explosive_plays %>%
  mutate(helmet = file.path("helmets", paste0(posteam, "_right.png")))

explosive_plays2 <- explosive_plays %>%
  arrange(desc(explosive_share))

# Plot with logos on y-xis
explosive_plays_plot_logos <- explosive_plays %>%
  ggplot(aes(x = explosive_share, y = reorder(helmet, explosive_share))) +
  geom_col(fill = if_else(explosive_plays$posteam == "SF", "#AA0000", "grey")) +
  # Share printed just past the end of each bar; the x expansion below makes
  # room for it inside the panel.
  geom_text(aes(label = scales::percent(explosive_share, accuracy = 0.1)),
            hjust = -0.15, family = "Titillium Web", size = 3, colour = "grey30") +
  labs(
    x = "Explosive Play Share (%)",
    y = "",
    caption = "Data: @nflscrapR | Plot: @steodosescu",
    title = glue("Most Explosive Teams, 2026 NFL Season"),
    subtitle = glue("Explosive plays = 10 yards+ on rushes, 20 yards+ on passes. As of Week {current_week}.")
  ) +
  theme_custom() +
  theme(plot.title = element_text(face = "bold", 
                                  size = 20, 
                                  hjust = 0.5
    )) +
  theme(plot.subtitle = element_markdown()) +
  theme(plot.title = element_markdown()) +
  geom_vline(xintercept = mean(explosive_plays$explosive_share), color = "red", linetype = "dashed") +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.1, suffix = " %"),
                     expand = expansion(mult = c(0, 0.08))) +
  theme(axis.text.y = nflplotR::element_path(size = 0.95)) +
  theme(legend.position = "none")

explosive_plays_plot_logos

```


***
## Offensive Series Success

A series ends when the offense gains a first down or the possession ends, so a single drive is usually made up of several series. Sorting by series success rate — the share of series ending in a touchdown or a first down — shows which offenses keep the chains moving rather than which simply score most often.

```{r Series Success, echo = FALSE, fig.height = 10}

# Series results, in the style of rbsdm.com's "Offensive Series Results".
# Uses `data` (all plays), not pbp_rp: pbp_rp is filtered to rushes/passes,
# which would drop the punt and field goal series entirely.
#
# One row per series, taken from the series' FIRST play. Do not substitute
# distinct(game_id, posteam, series, series_result): after a defensive
# touchdown the extra-point play carries the *scoring* team as posteam while
# keeping the same series number and series_result ("Opp touchdown"), so
# distinct() emits that series twice and charges the team that just scored
# with a failed offensive series.
series_levels <- c("TD", "First Down", "FG", "Punt", "TO")

series_results <- data %>%
  filter(season_type == "REG", !is.na(posteam), posteam != "", !is.na(series_result)) %>%
  group_by(game_id, series) %>%
  slice_min(play_id, with_ties = FALSE) %>%
  ungroup() %>%
  # Not genuine offensive attempts.
  filter(!series_result %in% c("QB kneel", "End of half")) %>%
  mutate(series_bucket = factor(case_when(
      series_result == "Touchdown"  ~ "TD",
      series_result == "First down" ~ "First Down",
      series_result == "Field goal" ~ "FG",
      series_result == "Punt"       ~ "Punt",
      # Turnover, Turnover on downs, Opp touchdown, Missed field goal.
      TRUE                          ~ "TO"
    ), levels = series_levels))

# .drop = FALSE keeps a zero row for buckets a team never reached, so every
# team keeps all five segments and the shares still sum to 1.
series_shares <- series_results %>%
  count(posteam, series_bucket, .drop = FALSE) %>%
  group_by(posteam) %>%
  mutate(n_series = sum(n),
         share = n / n_series,
         # Series success = touchdown or first down; this sorts the chart.
         success_rate = sum(share[series_bucket %in% c("TD", "First Down")])) %>%
  ungroup() %>%
  # Helmets rather than logos on the y axis. nflplotR::element_path() renders
  # axis text that is a file path as an image, so the axis values are the
  # paths themselves and the ordering rides along on the factor.
  mutate(helmet = file.path("helmets", paste0(posteam, "_right.png")),
         helmet = fct_reorder(helmet, success_rate),
         # Ink colour for the in-bar labels: light text on the dark fills,
         # dark text on the light ones. scale_colour_identity() below uses
         # these literally rather than mapping them through a scale.
         label_ink = case_when(
           series_bucket %in% c("TD", "FG", "TO") ~ "floralwhite",
           TRUE                                   ~ "grey15"
         )) %>%
  # Label positions computed explicitly rather than via position_stack().
  # Mapping colour in the text layer changes that layer's implicit grouping,
  # which made position_stack() order the labels differently from the bars
  # and parked each team's FG share near the left edge of its row.
  group_by(posteam) %>%
  arrange(series_bucket, .by_group = TRUE) %>%
  mutate(label_x = cumsum(share) - share / 2) %>%
  ungroup()

# One row per team for the abbreviations down the right-hand side.
series_team_labels <- series_shares %>%
  distinct(posteam, helmet)

# An ordered outcome scale, so diverging rather than categorical: blue = the
# offense kept the ball moving, red = it did not, neutral grey for FG in the
# middle. Stepped so that adjacent segments stay apart for colourblind
# readers (worst adjacent pair is deltaE 23.5 under deuteranopia).
series_cols <- c("TD" = "#08415F", "First Down" = "#6FB3DC", "FG" = "#6E665C",
                 "Punt" = "#E8A97E", "TO" = "#93261F")

series_results_plot <- series_shares %>%
  ggplot(aes(x = share, y = helmet, fill = series_bucket)) +
  # Drawing the segment borders in the panel colour leaves a hairline gap, so
  # neighbouring fills never touch.
  geom_col(width = 0.78, position = position_stack(reverse = TRUE),
           colour = "floralwhite", linewidth = 0.45) +
  # Segments under 5% are too narrow to hold a legible label, so they are
  # left blank rather than overplotted. Raise or lower the threshold here.
  geom_text(aes(x = label_x,
                label = if_else(share >= 0.05,
                                scales::percent(share, accuracy = 1), ""),
                colour = label_ink),
            family = "Titillium Web", size = 3, show.legend = FALSE) +
  scale_colour_identity() +
  # Team abbreviations down the right edge, inside the panel padding added by
  # the x expansion below (so no clip = "off" needed).
  geom_text(data = series_team_labels, aes(x = 1.015, y = helmet, label = posteam),
            inherit.aes = FALSE, hjust = 0, family = "Titillium Web",
            size = 3.2, fontface = "bold", colour = "grey30") +
  scale_fill_manual(values = series_cols, name = NULL) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.06))) +
  labs(
    x = "Series Result Share",
    y = "",
    caption = "Data: @nflfastR | Plot: @steodosescu",
    title = glue("Offensive Series Results, 2026 NFL Season"),
    subtitle = glue("How every offensive series ended. Sorted by series success rate (TD or first down). As of Week {current_week}.")
  ) +
  theme_custom() +
  theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5)) +
  theme(plot.subtitle = element_markdown(hjust = 0.5)) +
  theme(legend.position = "top") +
  theme(panel.grid.major.y = element_blank()) +
  theme(axis.text.y = nflplotR::element_path(size = 0.95))

series_results_plot



```

